---
title: "KIMOVE"
author: "Jonathan Tjerkaski"
date: "24 April 2019"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

```


# Aim

The aim of this assignment is to investigate of cortical thickness can be used  
to succesfully predict the correct diagnosis of patients with movement disorders. 

# Packages

##Install packages

This step only has to be done once!

```{r install_packages, eval = FALSE}

install.packages("tidymodels")
install.packages("tidyverse")
install.packages("knitr")
install.packages("AppliedPredictiveModeling")
install.packages("caret")
install.packages("MLmetrics")
install.packages("ggfortify")
install.packages("kableExtra")
install.packages("summarytools")
install.packages("readxl")
install.packages("corrr")
install.packages("RColorBrewer")
install.packages("corrplot")
#This may be required for the Support Vector Machines analysis
install.packages("kernlab")
install.packages('modelgrid')
install.packages('e1071')
install.packages('MASS')
install.packages('descr')

```

##Load Packages


```{r packages}

library(tidymodels)
library(tidyverse)
library(knitr)
library(AppliedPredictiveModeling)
library(caret)
library(MLmetrics)
library(ggfortify)
library(kableExtra)
library(summarytools)
library(corrplot)
library(readxl)
library(corrr)
library(modelgrid)
library(e1071)
library(MASS) 
library(descr) 


options(knitr.kable.NA = '')


```

#The data

```{r data}

data <- read_excel('datatabell_190424_final.xlsx')

```


## Pre-process the data table

```{r pre_process_data}

data <-  data %>% 
  select(KIMOVE_ID:Model) %>% 
  select( -("VSGP@MRI":"ProbMSA@MRI")) %>% 
  rename_all(tolower) %>% 
  rename("gender" = "gender (0=male,1=female)", "h_y" = "h&y", "symptom_duration"
         = "symtomdur", "diagnosis" = "diagnos") %>% 
  as_tibble()

```

###changes characters to numerics and vice-versa

Change the thicknesses to numerical variables, as they are currently falsely 
believed to be character strings. age, symptom_duration also. diagnosis and 
gender should be factors, but are currently numeric. As only Kimove_id and h_y 
are correctly labeles as character and numeric, respectively, the simplest is to
just change all numerics into characters and vice versa and then return kimove_id 
to characted and h-y to numeric. However, if I change kimove_id to numeric it 
will be lost, so I will not select it when transforming the variables.

```{r factors_numerics}

Kimove_id <- data$kimove_id

Model <- data$model

data <- data %>%
  mutate_if(is.character, as.numeric) %>% 
  mutate(gender = as.factor(gender),
         diagnosis = as.factor(diagnosis)) 

data$kimove_id <- Kimove_id

data$model <- Model

```

###change factor levels of columns

"diagnosis" from "1,2,3" to "PD,PSP,MSA" and "gender" from "0" and "1" to "male"
and "female"

```{r factor_levels}

data <- data %>%
  mutate(diagnosis = fct_recode(diagnosis,
    "PD"  = "1", 
    "PSP" = "2",
    "MSA" = "3")) %>% 
  mutate(gender = fct_recode(gender,
    "male"  = "0",
    "female" = "1"))
  
```

# Descriptive statistics and correlations

##Demographic

```{r demographic, results = 'asis', comment = NA, prompt = FALSE, cache = FALSE, eval = FALSE}

demography <- data %>% 
  select(diagnosis, age, gender, h_y, symptom_duration, model) %>% 
  rename("Hoehn and Yahr scale" = h_y, "Symptom duration" = symptom_duration)

print(dfSummary(demography, plain.ascii = FALSE, style = "grid", 
          graph.magnif = 0.75, valid.col = FALSE, na.col = FALSE, varnumbers = F,
          tmp.img.dir = "/tmp"))
  
 print(stby(data = demography, INDICES = data$diagnosis, FUN = descr,
           stats = c("mean", "sd", "min", "med", "max"),  transpose = TRUE))

```



##Shapiro-wilk test for normality

```{r normality_test}

 data %>% 
    keep(is.numeric) %>% 
    select(-h_y) %>% 
    gather(key = "variable_name", value = "value") %>% 
    group_by(variable_name)  %>% 
    do(tidy(shapiro.test(.$value))) %>% 
    ungroup() %>% 
    select(-method) %>% 
    kable(digits = 3, caption = "Shapiro-Wilk test for normality of all numeric variables") 

# sw_test_results <- data %>% 
#     gather(key = "variable_name", value = "value", c("age", "symptom_duration", "lh_bankssts_thickness":"etiv")) %>% 
#     group_by(variable_name, diagnosis)  %>% 
#     do(tidy(shapiro.test(.$value))) %>% 
#     ungroup() %>% 
#     arrange(diagnosis)


# kable(sw_test_results, digits = 3, format = "latex", longtable = T, caption =
#         "Shapiro-Wilk test for normality by diagnosis")  


```


## $\chi^2$ - test

$$\chi^2 = \sum \frac {(O - E)^2}{E}$$
Diagnosis is the response variable and gender is the explanatory variable

```{r chisq}


data %>% 
  chisq_test(formula = diagnosis ~ gender) %>% 
  rename("Chisq-statistic" = statistic,"Degrees of freedom" = chisq_df,
         "p-value" = p_value) %>% 
  kable(digits = 3, caption = "chisq-test")

data %>% 
  chisq_test(formula = diagnosis ~ model) %>% 
  rename("Chisq-statistic" = statistic,"Degrees of freedom" = chisq_df,
         "p-value" = p_value) %>% 
  kable(digits = 3, caption = "chisq-test")

CrossTable(data$diagnosis, data$model,
       fisher = T, chisq = T, expected = T,
       prop.c = F, prop.t = F, prop.chisq = F, 
       sresid = T, format = 'SPSS')

```

## One-way Analysis of variance (ANOVA)

Age as the dependant variable and diagnosis as the independant variable

```{r anova}

anovar <- aov(age ~ diagnosis, data = data) 

anovar %>% 
  tidy() %>% 
kable( digits = 3, caption = "One-way ANOVA")

tidy(TukeyHSD(anovar, which = 'diagnosis')) %>% 
  kable(digits = 3, caption = "Tukey post-hoc test")

```


##Kruskall-Wallis test

```{r KW}

kw <- kruskal.test(age ~ diagnosis, data = data)

kw %>% 
  tidy() %>% 
kable( digits = 3, caption = "Kruskall-Wallis test")

tidy(pairwise.wilcox.test(data$age, data$diagnosis, p.adj = "bonf")) %>% 
  kable(digits = 3, caption = "Pair-wise Wilcoxon rank-sum test")

```


##Skewness of numeric variables.

```{r skew}

data %>%
  summarise_if( is.numeric, e1071::skewness ) %>%
  kable(digits = 3, caption = "Skewness")

```

##Age distribution

Age was found not to be normally distributed by the Shapiro-Wilk test, and these 
plots sort of confirm that notion. The age variable appears to have a slightly
skewed distribution.

```{r age_distribution}

ggplot(data, aes(x=age))+
  geom_histogram(aes(y=..density..), alpha=0.5, 
                position="identity", color="darkblue") + 
  geom_density(color="darkblue", alpha=.7)+
geom_vline(aes(xintercept=mean(age)),
            color="blue", linetype="dashed", size=1, alpha = .4)+
  geom_text(aes(x=mean(age), label="Mean", y=0), colour="red", 
        vjust = 1.2, text=element_text(size=14)) +
  ggtitle("Age - Histogram")+
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data, aes(x=age, fill = diagnosis))+
  geom_density(color="darkblue") + 
  ggtitle("Age by diagnosis") +
  theme(plot.title = element_text(hjust = 0.5))


```


##Pearson correlation coefficients between brain regions

A note on the interpretation of these tables: The purpose of obtaining the 
correlation coefficients was to see if several regions were highly correlated.
This information may be of value when choosing machine learning algorithm, as 
some are not compatible with higly correlated data. These correlation
coefficients were made on the entire data and not for each respective diagnosis.

```{r correlation}


data %>%
  select("lh_bankssts_thickness" : "rh_meanthickness_thickness", -(c("rh_meanthickness_thickness","lh_meanthickness_thickness" ))) %>%
  correlate() %>%
  shave() %>% 
  stretch(na.rm = TRUE) %>% 
  filter(r > 0.8) %>% 
  mutate(x = str_replace(string=x, pattern='_thickness', replacement="")) %>% 
  mutate(y = str_replace(string=y, pattern='_thickness', replacement="")) %>%
  arrange(desc(r)) %>% 
  kable(digits = 3, caption = "Brain regions with the greatest correlation")

data %>%
  select("lh_bankssts_thickness" : "rh_meanthickness_thickness", -(c("rh_meanthickness_thickness","lh_meanthickness_thickness" ))) %>%
  correlate() %>%
  shave() %>% 
  stretch(na.rm = TRUE) %>%
  top_n(-10) %>%
  mutate(x = str_replace(string=x, pattern='_thickness', replacement="")) %>% 
   mutate(y = str_replace(string=y, pattern='_thickness', replacement="")) %>% 
  arrange(r) %>% 
  kable(digits = 2, caption = "The 10 least correlated brain regions")
 
  
data %>%
  select("lh_bankssts_thickness" : "rh_meanthickness_thickness", -(contains("meanthickness"))) %>%
  correlate() %>%
  shave() %>% 
  stretch(na.rm = TRUE) %>% 
  filter(str_detect(x, "^lh"),
         str_detect(y, "^rh")) %>% 
  mutate(x = str_replace(string=x, pattern='lh_', replacement="")) %>% 
  mutate(y = str_replace(string=y, pattern='rh_', replacement="")) %>% 
  filter(x == y) %>% 
  select("Region" = y, r) %>%
  mutate(Region = str_replace(string=Region, pattern='_thickness',      replacement="")) %>% 
  arrange(desc(r)) %>% 
  kable(digits = 2, caption = "Correlation between the left and the right hemisphere")
  
  

```

###Correlation plot

Heatmap of the correlations, ordered using hierarchical clustering.

```{r corrplot, results='asis'}

data %>%
  select(age:etiv, -gender) %>% 
  cor(.) %>%
  corrplot(., order = "hclust", tl.cex = .35)

```

From the graph, it seems that there are groups of predictors, that have strong positive correlations (dark blue).

#PCA

```{r pca}

kimove_pca <- data %>% 
  select(-kimove_id, -model) %>% 
  mutate(gender = ifelse(gender == "male", 0, 1)) %>% 
  nest() %>% 
  mutate(pca = map(data, ~prcomp(.x %>% select(-diagnosis), center = TRUE, scale = TRUE)), 
         pca_aug = map2(pca, data, ~broom::augment(.x, data = .y))) 

```

##results

```{r pca_results}

var_exp <- kimove_pca %>%
  unnest(pca_aug) %>% 
  summarize_at(.vars = vars(contains("PC")), .funs = funs(var)) %>% 
  gather(key = pc, value = variance) %>% 
  mutate(var_exp = variance/sum(variance),
         cum_var_exp = cumsum(var_exp),
         pc = str_replace(pc, ".fitted", ""))

```
## Variance explained by each of the 9 first principal components

```{r var_exp}

var_exp %>% 
  slice(1:9) %>% 
  rename(
    `Variance Explained` = var_exp,
    `Cumulative Variance Explained` = cum_var_exp
  ) %>% 
  gather(key = key, value = value, `Variance Explained`:`Cumulative Variance Explained`) %>% 
  ggplot(aes(pc, value, group = key)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~key, scales = "free_y") +
  theme_bw() +
  lims(y = c(0, 1)) +
  labs(y = "Variance",
       title = "Variance explained by the 9 first principal component",
       subtitle = "Note: There are 76 PC's in total")
       

```

##Screeplot

A barplot showing the variances for each Principal Component

```{r pca_plot}

var_exp %>% 
mutate(var_exp = var_exp*100) %>% 
  slice(1:9) %>% 
  arrange(var_exp) %>% 
ggplot( aes(x=pc, y = var_exp, fill = pc)) +
  geom_bar(stat="identity") + 
  geom_text(aes(label= round(var_exp)), vjust=1.2, color="white", size=3.5) +
  xlab("Principal Components") +
  ylab("Variance Explained") +
  theme_bw() +
  ggtitle("Scree plot of the 9 first principle components (out of a total of 76 PC's)")



```


##PCA graph

autoplot.prcomp() can take any arguments that can be passed to ggbiplot(), 
so to see all of your options use ?ggbiplot.

```{r pca_graph}

var_per <- var_exp %>% 
mutate(var_per = round(var_exp*100)) %>%
  select(var_per)

kimove_pca %>% 
   mutate(
    pca_graph = map2(
      .x = pca,
      .y = data,
      ~ autoplot(.x, data = .y, colour = "diagnosis", frame = TRUE,
                 frame.type = 'norm') +
        theme_bw() +
        labs(x = paste("PC1", "(", var_per[ 1, 1], "%", ")", sep=""),
             y = paste("PC2 - ", var_per[ 2, 1], "%", sep=""),
             title = "First two principal components of PCA on the 
             kimove dataset")
    )
  ) %>%
  pull(pca_graph)



```

PCA was not very usefull here!

#Support Vector Machine

##split the data

Randomly split the data into a training dataset and a test dataset using
approximately 75% of the data in the training dataset, with the remaining 
approximately 25% ending up in the test dataset. The software is also instructed
to maintain the same distribution of the diagnoses as is in the original data, 
so that e.g. 75% of PSP patients are allocated to  the the training dataset, 
which constitute the same proportion of the total patients as PSP does from the 
original, unsplit, data.


```{r split}

kimove <- data %>% 
  select(-c(kimove_id, model)) 
  

set.seed(123)

train_test_split <- initial_split(kimove, strata = "diagnosis", prop = .75)

diagnosis_train <- training(train_test_split)
diagnosis_test <- testing(train_test_split)

```


##Create recipes in order to pre-process the data

step_dummy creates dummy variable e.g. turns male and female in gender into 
two different columns that have 1's and 0's. The number of new columns created 
is by default the number of different labels in a column (e.g. male and female) 
minus one. Thus, for gender only 1 column exists and no new column is created.

step_scale och step_center scale all variables equally, with a mean of 0 and 
a standard deviation equal to 1. This is accomplished by subtracting the mean 
and dividing by the standard deviation.

step_boxcox performs boxcox transformation on all numeric variables. This can 
involve doing a logarithmic transformation of one variable and taking the squared 
root of all rows in a column, depending on a certain lamda value which is 
calculated for each column. This makes the data more normaly distributed.
The Box-Cox transformation, which requires a strictly positive variable, can thus 
be used to rescale a variable to be more similar to a normal distribution. 
In this package, the partial log-likelihood function is directly 
optimized within a reasonable set of transformation values 
(which can be changed by the user).

This transformation is typically done on the outcome variable using the 
residuals for a statistical model (such as ordinary least squares). 
Here, a simple null model (intercept only) is used to apply the transformation 
to the predictor variables individually. 
This can have the effect of making the variable distributions more symmetric.

step_pca extracts the principle components of the data and then replaces the 
numeric variables with the principle components.Threshold set att 9% of the 
variance explained, corresponding to the first 61 principle components out of 
a total of 76 principle components for this data (see the object 
var_exp from the pca above)

step_downsample: The default value for the ratio argument (1) is used, which 
means that all other levels are sampled down to have the same frequency as the 
least occurring level. A value of 2 would mean that the majority levels will 
have (at most) (approximately) twice as many rows than the minority level.

step_corr potentially removes variables that have large absolute correlations 
(above the specified threshold) with other variables. The default measure for 
correlation is the Pearson correlation coefficient, which is used for parametric 
data. The data was evaluated above for normality using the shapiro wilk test.
It is highly questionable that the data is normally distributed in any way, as
multiple variables had low p-values in the shapiro wilk test. 
However, for the purposeof this analysis it does not matter, as the 
coefficient of variation is simply used as a means of filtering data. An 
alternative means of filtering using step-corr could for instance be using
Spearman's coefficient of correlation, which is generally considered better than
Pearson's for non-parametric data. However, as the MRI data underwent smoothing,
it is assumed that the Pearsons coefficient is also appropriate for use on this 
data.

step_nzv removes variables with zero variance and "near-zero" variance. 
These variables either have very few unique values relative to the number 
of samples and/or the ratio of the frequency of the most common value to the 
frequency of the second most common value is large. Variables are determined to 
be near-zero variables if The ratio of the most common value to the second most 
common value is less than 95:5 or if The percentage of distinct values out of 
the number of total samples is less than 10


```{r recipe}

rec_orig <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
step_dummy(all_nominal(), - all_outcomes())

rec_obj <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
step_dummy(all_nominal(), - all_outcomes()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric())

rec_boxcox <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
step_BoxCox(all_numeric()) %>%
step_dummy(all_nominal(), - all_outcomes()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric())

rec_pca <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
step_dummy(all_nominal(), - all_outcomes()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric()) %>%
step_pca(all_numeric(), threshold = .99)

rec_both <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
step_BoxCox(all_numeric()) %>%
step_dummy(all_nominal(), - all_outcomes()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric()) %>%
step_pca(all_numeric(), threshold = .99)

rec_corr <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
step_corr(all_numeric(), threshold = .8) %>%
step_dummy(all_nominal(), - all_outcomes()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric())
  
rec_down <-  recipe(diagnosis ~ ., data = diagnosis_train) %>%
step_dummy(all_nominal(), - all_outcomes()) %>% 
step_center(all_numeric()) %>%
step_scale(all_numeric()) %>%
step_downsample(diagnosis)

#The following models were tested during analysis, but did not work well!
# rec_corr_down <-  recipe(diagnosis ~ ., data = diagnosis_train) %>%
# step_corr(all_numeric(), threshold = .80) %>%
# step_dummy(all_nominal(), - all_outcomes()) %>% 
# step_center(all_numeric()) %>%
# step_scale(all_numeric()) %>%
# step_downsample(diagnosis)
# 
# rec_corr_pca <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
# step_corr(all_numeric(), threshold = .80) %>%
# step_dummy(all_nominal(), - all_outcomes()) %>%
# step_center(all_numeric()) %>%
# step_scale(all_numeric()) %>%
# step_pca(all_numeric(), threshold = .99)
# 
# rec_boxcor <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
# step_BoxCox(all_numeric()) %>%
# step_dummy(all_nominal(), - all_outcomes()) %>% 
# step_corr(all_numeric(), threshold = .80) %>%
# step_center(all_numeric()) %>%
# step_scale(all_numeric())  
#
# rec_nzv <-recipe(diagnosis ~ ., data = diagnosis_train) %>%
# step_nzv(all_predictors()) %>%
# step_dummy(all_nominal(), - all_outcomes()) %>%
# step_center(all_numeric()) %>%
# step_scale(all_numeric())
 


```


###Metric F-score

$$F  = 2 \times \frac{Precision \times Recall}{Precision + Recall}$$

F-score is defined as the harmonic mean of Precision and Recall. F-score is
a better metric of model performance when using unbalanced data 
(i.e. many more PD patients than patients with MSA or PSP) than simply the
proportion of correctly classified instances, also known as "Accuracy" 
in the context of Machine Learning
$(Accuracy=\frac{Correct\: Predictions}{Total\: number\: of \:Predictions})$.

This is due to the fact that F-score takes into account both Precision and 
Recall (Sensitivity). For instance, If a model is good at predicting PD patients
but bad at predicting PSP and MSA, the accuracy will be very high, but the model
does not perform in such a way which we would consider to be clinically useful. 
By using both precision and recall, F-score accounts for not only the 
true positives, but also the number of false negatives that in the example above 
would help in showing that the models performance is, in fact, not quite as good 
as the Accuracy would make it seem.


```{r f_score}

F1 <- function(data, lev = NULL, model = NULL) {
  f1_val <- F1_Score(y_pred = data$pred, y_true = data$obs, positive = lev[1])
  c(F1 = f1_val)
}


```


###Preparation for 10-fold cross-validation

Training settings for models using the kappa metric below.

```{r train_control_kappa}

control <- trainControl(
 method = "repeatedcv",
 number = 10,
 repeats = 10)

```

Training settings for models using F-score

```{r train_control_f_score}

ctrl <- trainControl(
 method = "repeatedcv",
 number = 10,
 repeats = 10,
 summaryFunction = F1, classProbs = TRUE)


```


## Choosing threshold 

##step_corr thershold

One way of dealing with the potential drawbacks of the observed "collinearity clusters" 
found in the correlation plot above is to apply a correlation filter. The 
correlation filter poses a heuristic approach to dealing with highly correlated 
predictors. It removes the predictors with the highest between-predictor 
correlations one at a time, until all between-predictor correlations are below 
some critical threshold.

modelgrid works by first creating an empty model grid with the model_grid() 
function. Then share_settings() sets shared settings, that will apply to all 
models by default. Then the models are trained using Caret package train function 
and the statistics can be displayed using Caret package.


Note: The following code was not run when creating this document, it was 
only used during the analysis to aid in choosing an appropriate threshold for 
step_corr


```{r corr_threshold, eval = FALSE}


models_cor <- 
  model_grid() %>%
  share_settings(
    data = diagnosis_train,
    trControl = control,
    metric = "Kappa",
    method = "svmRadial",
    tuneLength = 12
    )

initial_recipe <- rec_obj

models_cor <- models_cor %>%
  add_model(model_name = "baseline", 
            x = initial_recipe)

models_cor <- models_cor %>%
  add_model(model_name = "corr_.7", 
            x = initial_recipe %>%
              step_corr(all_predictors(), threshold = .7)) %>%
  add_model(model_name = "corr_.75", 
            x = initial_recipe %>%
              step_corr(all_predictors(), threshold = .75)) %>%
  add_model(model_name = "corr_.8", 
            x = initial_recipe %>%
              step_corr(all_predictors(), threshold = .8)) %>%
  add_model(model_name = "corr_.85", 
            x = initial_recipe %>%
              step_corr(all_predictors(), threshold = .85)) %>%
  add_model(model_name = "corr_.9", 
            x = initial_recipe %>%
              step_corr(all_predictors(), threshold = .9))%>%
  add_model(model_name = "corr_.95", 
            x = initial_recipe %>%
              step_corr(all_predictors(), threshold = .95))

models_cor <- models_cor %>% train(.)

models_cor$model_fits %>%
  map(pluck(c("recipe", "term_info", "role"))) %>%
  map_int(~ sum(.x == "predictor")) %>% 
kable(digits = 3, caption = "Number of columns kept after the step_corr function")

#resamps_cor <- models_cor$model_fits %>% resamples(.)

#summary(resamps_cor)


```


##step_pca threshold

Another approach to dealing with highly correlated predictors is to apply a
Principal Component Analysis transformation of the predictors in order to reduce
the dimensions of data set.


Note: The following code was not run when creating this document, it was 
only used during the analysis to aid in choosing an appropriate threshold for 
step_pca* 
 

```{r pca_threshold, eval = FALSE}

models_pca <- 
  model_grid() %>%
  share_settings(
    data = diagnosis_train,
    trControl = control,
    metric = "Kappa",
    method = "svmRadial",
    tuneLength = 12
    )

models_pca <- models_pca %>%
  add_model(model_name = "baseline", 
            x = initial_recipe)

models_pca <- models_pca %>%
  add_model(model_name = "pca_.75", 
            x = initial_recipe %>%
              step_pca(all_predictors(), threshold = .75)) %>%
  add_model(model_name = "pca_.8", 
            x = initial_recipe %>%
              step_pca(all_predictors(), threshold = .8)) %>%
  add_model(model_name = "pca_.85",
            x = initial_recipe %>%
              step_pca(all_predictors(), threshold = .85)) %>%
  add_model(model_name = "pca_.9", 
            x = initial_recipe %>%
              step_pca(all_predictors(), threshold = .9)) %>%
  add_model(model_name = "pca_.95",
            x = initial_recipe %>%
              step_pca(all_predictors(), threshold = .95)) %>%
  add_model(model_name = "pca_.95",
            x = initial_recipe %>%
              step_pca(all_predictors(), threshold = .99))


models_pca <- train(models_pca)

models_pca$model_fits %>%
  map(pluck(c("recipe", "term_info", "role"))) %>%
  map_int(~ sum(.x == "predictor")) %>% 
kable(digits = 3, caption = "Number of PC's kept after the step_pca function")

#resamps_pca <- models_pca$model_fits %>% resamples(.)

#summary(resamps_pca)

```


###check variables removed by step_cor

Using a threshold of Pearson correlation coefficient greater than 0.8.

```{r check_cor}

check_cor <- prep(rec_corr, training = diagnosis_train)


kable(check_cor$steps[[1]]$removals, col.names = "Removed variables", 
caption = "step_corr removals")


```

Comment: many of the regions removed by step_corr due to high correlation are 
also the ones that were found to have statistically significant cortical 
thinning in QDEC.


###check number of principle components used by step_pca

Using a threshold of 99% of the variance explained.

```{r check_pca}

check_pca <- prep(rec_pca, training = diagnosis_train)


kable(check_pca$steps[[4]]$num_comp, col.names = "Principle Components", 
caption = "step_pca Principle components retained")

```

###check lambda values calculated by step_boxcox


Note: The following code was not run when creating this document, it took
to much space*


```{r lambda}

check_boxcox <- prep(rec_boxcox, training = diagnosis_train)


kable(check_boxcox$steps[[1]]$lambdas, col.names = "Lambda", 
caption = "step_boxcox Lambda values")

```


#Train models using the "Caret" package in R

SVM using a (Gaussian) radial basis function kernel has two hyperparameters: 
Cost (also just simply referred to as "C") and $\gamma$ or $\sigma$. $\gamma$
and $\sigma$ are related and can be used interchangeably, as shown by the 
following equation:

$$\gamma = \frac{1}{2\sigma^2}$$
The number of Cost parameters that are tested during each crossvalidation step
are denoted by the Tunelength paramater in the "train" function from the "Caret"
package.

$\sigma$ is calculated by the "sigest" function in the kernlab pakage, which is 
called upon by the "train" function from "Caret". the optimal values of the 
hyper-parameter of sigma are shown to lie in between the 0.1 and 0.9 quantile of 
the $||x-x'||$ statistic. The "sigest" from the "kernlab" package function 
uses a sample of the training set to estimate the quantiles and returns a vector 
containing the values of these quantiles. It has empirically been shown by Caputo,
Sim, Furesjo, and Smola in an article from 2002 that any value within this interval 
leads to a good performance in the model. A value from this interval is then used
the value for $\sigma$ by the "train" function from "Caret" and it is kept 
constant as long as the training data which is used is the same. For instance, 
the value for $\sigma$ is slightly different in the model which uses a 
downsampled dataset than that which is used in the other models.

```{r svm_f_score}

set.seed(123)

svm <- train(rec_obj, data = diagnosis_train,
method = "svmRadial",
tuneLength = 12,
trControl = ctrl)

set.seed(123)

boxcox <- train(rec_boxcox, data = diagnosis_train,
method = "svmRadial",
tuneLength = 12,
trControl = ctrl)

set.seed(123)

pca <- train(rec_pca, data = diagnosis_train,
method = "svmRadial",
tuneLength = 12,
trControl = ctrl)

set.seed(123)

both<- train(rec_both, data = diagnosis_train,
method = "svmRadial",
tuneLength = 12,
trControl = ctrl)

set.seed(123)

orig <- train(rec_orig, data = diagnosis_train,
method = "svmRadial",
tuneLength = 12,
trControl = ctrl)

set.seed(123)

svm_corr <- train(rec_corr, data = diagnosis_train,
method = "svmRadial",
tuneLength = 12,
trControl = ctrl)

set.seed(123)

down <- train(rec_down, data = diagnosis_train,
method = "svmRadial",
tuneLength = 12,
trControl = ctrl)

models <- list("SVM" = svm, "BoxCox" = boxcox, "PCA" = pca, "Both" = both, "Orig" = orig, "SVM_corr" = svm_corr, "Down" = down )

```


##Results from the 10-fold cross validation?

```{r resample_result_f_score}

resamps <- resamples(models)

print(summary(resamps))


```

##Model performance using the "Yardstick" - package


```{r svm_f_score_model_performance}

testing_results <- diagnosis_test %>%
        mutate("SVM" = predict(svm, diagnosis_test)) %>% 
        mutate("BoxCox" = predict(boxcox, diagnosis_test)) %>% 
        mutate("PCA" = predict(pca, diagnosis_test)) %>% 
        mutate("Both" = predict(both, diagnosis_test)) %>% 
        mutate("Orig" = predict(orig, diagnosis_test)) %>%  
        mutate("SVM_corr" = predict(svm_corr, diagnosis_test)) %>% 
        mutate("Down" = predict(down, diagnosis_test))
  
  
SVM <- conf_mat(testing_results, truth = diagnosis, estimate = SVM)

autoplot(SVM, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM") +
  theme(plot.title = element_text(hjust = 0.5))

SVM <- (summary(SVM)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(SVM,digits = 3, caption = "Support Vector Machine")


PCA <- conf_mat(testing_results, truth = diagnosis, estimate = PCA)

autoplot(PCA, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM with PCA") +
  theme(plot.title = element_text(hjust = 0.5))

PCA <- (summary(PCA)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(PCA,digits = 3, caption = "SVM with PCA")

BOXCOX <- conf_mat(testing_results, truth = diagnosis, estimate = BoxCox)

autoplot(BOXCOX, type = "heatmap") +
  ggtitle("Diagnosis Heatmap BoxCox") +
  theme(plot.title = element_text(hjust = 0.5))

BOXCOX <- (summary(BOXCOX)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(BOXCOX,digits = 3, caption = "SVM with BoxCox")

Both <- conf_mat(testing_results, truth = diagnosis, estimate = Both)

autoplot(Both, type = "heatmap") +
  ggtitle("Diagnosis Heatmap with both boxcox and pca") +
  theme(plot.title = element_text(hjust = 0.5))

Both <- (summary(Both)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(Both,digits = 3, caption = "SVM with both BoxCox and PCA")

Orig <- conf_mat(testing_results, truth = diagnosis, estimate = Orig)

autoplot(Orig, type = "heatmap") +
  ggtitle("Diagnosis Heatmap svm without centering and scaling") +
  theme(plot.title = element_text(hjust = 0.5))

Orig <- (summary(Orig)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(Orig,digits = 3, caption = "SVM without centering and scaling")

SVM_corr <- conf_mat(testing_results, truth = diagnosis, estimate = SVM_corr)

autoplot(SVM_corr, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM without correlation") +
  theme(plot.title = element_text(hjust = 0.5))

SVM_corr <- (summary(SVM_corr)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(SVM_corr,digits = 3, caption = "SVM without correlations")

Down <- conf_mat(testing_results, truth = diagnosis, estimate = Down)

autoplot(Down, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM with downsampling") +
  theme(plot.title = element_text(hjust = 0.5))

Down <- (summary(Down)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(Down,digits = 3, caption = "Support Vector Machine with downsampling")

```


## Model performance for each diagnosis

```{r svm_f_score_model_performance_per_diagnosis}

SVM_cf <- tidy(confusionMatrix(testing_results$diagnosis,testing_results$SVM)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(SVM_cf, digits = 3, caption = "Support Vector Machine")

Orig_cf <- tidy(confusionMatrix(testing_results$diagnosis,testing_results$Orig)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(Orig_cf, digits = 3, caption = "Support Vector Machine without scaling and centering")

boxcox_cf <- tidy(confusionMatrix(testing_results$diagnosis,testing_results$BoxCox)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(boxcox_cf, digits = 3, caption = "Support Vector Machine with BoxCox transformation")

pca_cf <- tidy(confusionMatrix(testing_results$diagnosis,testing_results$PCA)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(pca_cf, digits = 3, caption = "Support Vector Machine with PCA")

both_cf <- tidy(confusionMatrix(testing_results$diagnosis,testing_results$Both)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(both_cf, digits = 3, caption = "Support Vector Machine with both PCA and BoxCox transformation")

svm_corr_cf <- tidy(confusionMatrix(testing_results$diagnosis,
                                    testing_results$SVM_corr)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(svm_corr_cf, digits = 3, caption = "Support Vector Machine without correlation")

down_cf <- tidy(confusionMatrix(testing_results$diagnosis,testing_results$Down)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(down_cf, digits = 3, caption = "Support Vector Machine with downsampling")

```

#SVM using the metric "kappa" instead of F-score


##Train models using metric = "Kappa"

Like F-score, which considers both Precision and Recall, as well as 
Youden's index which takes into consideration both sensitivity and specificity, 
Kappa or "Cohen's Kappa" or simply $\kappa$ is a better metric of model 
performance than simply the proportion of correctly classified instances
if unbalanced data is uses, as κ takes into account not only the number of 
correct predictions, but also the possibility of the agreement occurring by chance.

$$\kappa = \frac{p_o-p_e}{1-p_e}$$ 
Where $p_o$ is the propportion of correct predictions out of the total number of
individual diagnoses included in this study (this is identical to "Accuracy" 
in the context of Machine Learning,
$Accuracy=\frac{Correct\: Predictions}{Total\: number\: of \:Predictions}$). 
$p_e$ is is the hypothetical probability of chance agreement between observed 
and predicted data, using the observed data to calculate the probabilities 
of  randomly observing the predicted results of  each category. 
If the predicted match the observed perfectly, $ {\kappa =1}$. If 
there is no concordance between the observed and predicted data, other than 
what would be expected by chance (as given by $p_e$), ${\kappa =0}$. Also, 
$\kappa$ can be negative, which that the model performs worse than what can be 
expected by random chance alone.

$p_e$ is calculated by the following equation:

$$p_e=\frac{1}{N^2}\sum_{k} n_{k1}n_{k2}$$   

As Cohen's Kappa was initially devised for use in inter-rater variability,
the variables in the equation for $p_e$ are described as follows:
"For categories k, number of items N and $n_{ki}$ is the number of times 
rater i predicted category k. The two raters are analogous to observed and 
predicted results in Machine learning.

```{r train_kappa, message = FALSE, warning = FALSE}

set.seed(123)

svm_k <- train(rec_obj, data = diagnosis_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

orig_k <- train(rec_orig, data = diagnosis_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

boxcox_k <- train(rec_boxcox, data = diagnosis_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

pca_k <- train(rec_pca, data = diagnosis_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

both_k <- train(rec_both, data = diagnosis_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

corr_k <- train(rec_corr, data = diagnosis_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)


set.seed(123)

down_k <- train(rec_down, data = diagnosis_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

models_kappa <- list("SVM" = svm_k, "BoxCox" = boxcox_k, "PCA" = pca_k, "Both" = both_k, "Orig" = orig_k, "Corr" = corr_k, "Down" = down_k)

```



##Results from the 10-fold cross validation on the models with metric = "Kappa"

```{r resamples_kappa}

resamps_kappa <- resamples(models_kappa)

print(summary(resamps_kappa))


```


##Model performance using metric = "Kappa" with the "Yardstick" - package

Same as above, but using metric = kappa to choose the optimal model

```{r svm_model_performance_kappa}

testing_results_kappa <- diagnosis_test %>%
        mutate("SVM" = predict(svm_k, diagnosis_test)) %>% 
        mutate("BoxCox" = predict(boxcox_k, diagnosis_test)) %>% 
        mutate("PCA" = predict(pca_k, diagnosis_test)) %>% 
        mutate("Both" = predict(both_k, diagnosis_test)) %>% 
        mutate("Orig" = predict(orig_k, diagnosis_test)) %>%  
        mutate("SVM_corr" = predict(corr_k, diagnosis_test)) %>% 
        mutate('Down' = predict(down_k, diagnosis_test)) 
        

SVM_k <- conf_mat(testing_results_kappa, truth = diagnosis, estimate = SVM)

autoplot(SVM_k, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM using Kappa metric") +
  theme(plot.title = element_text(hjust = 0.5))

SVM_k <- (summary(SVM_k)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(SVM_k,digits = 3, caption = "Support Vector Machine using Kappa metric") 



PCA_k <- conf_mat(testing_results_kappa, truth = diagnosis, estimate = PCA)

autoplot(PCA_k, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM with PCA using Kappa metric") +
  theme(plot.title = element_text(hjust = 0.5))

PCA_k <- (summary(PCA_k)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(PCA_k,digits = 3, caption = "SVM with PCA using Kappa metric")

BOXCOX_k <- conf_mat(testing_results_kappa, truth = diagnosis, estimate = BoxCox)

autoplot(BOXCOX_k, type = "heatmap") +
  ggtitle("Diagnosis Heatmap BoxCox using Kappa metric") +
  theme(plot.title = element_text(hjust = 0.5))

BOXCOX_k <- (summary(BOXCOX_k)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(BOXCOX_k,digits = 3, caption = "SVM with BoxCox using Kappa metric")

Both_k <- conf_mat(testing_results_kappa, truth = diagnosis, estimate = Both)

autoplot(Both_k, type = "heatmap") +
  ggtitle("Diagnosis Heatmap with both boxcox and pca using Kappa metric") +
  theme(plot.title = element_text(hjust = 0.5))

Both_k <- (summary(Both_k)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(Both_k,digits = 3, caption = "SVM with both BoxCox and PCA using Kappa metric")

Orig_k <- conf_mat(testing_results_kappa, truth = diagnosis, estimate = Orig)

autoplot(Orig_k, type = "heatmap") +
  ggtitle("Diagnosis Heatmap svm without centering and scaling using Kappa metric") +
  theme(plot.title = element_text(hjust = 0.5))

Orig_k <- (summary(Orig_k)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(Orig_k,digits = 3, caption = "SVM without centering and scaling using Kappa metric")

SVM_corr_k <- conf_mat(testing_results_kappa, truth = diagnosis, estimate = SVM_corr)

autoplot(SVM_corr_k, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM without correlation using Kappa metric") +
  theme(plot.title = element_text(hjust = 0.5))

SVM_corr_k <- (summary(SVM_corr_k)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(SVM_corr_k,digits = 3, caption = "SVM without correlations using Kappa metric")



Down_k <- conf_mat(testing_results_kappa, truth = diagnosis, estimate = Down)

autoplot(Down_k, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM with downsampling using Kappa metric") +
  theme(plot.title = element_text(hjust = 0.5))

Down_k <- (summary(Down_k)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(Down_k,digits = 3, caption = "Support Vector Machine with downsampling using Kappa metric")


```

Best performance by SVM with step_corr, followed by SVM with step_pca. 


## Model performance for each diagnosis using metric = "Kappa"

```{r svm_model_performance_kappa_by_diagnosis}

SVM_cf_kappa <- tidy(confusionMatrix(testing_results_kappa$diagnosis,testing_results_kappa$SVM)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(SVM_cf_kappa, digits = 3, caption = "Support Vector Machine using metric = kappa")

Orig_cf_kappa <- tidy(confusionMatrix(testing_results_kappa$diagnosis,testing_results_kappa$Orig)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(Orig_cf_kappa, digits = 3, caption = "Support Vector Machine without scaling and centering using metric = kappa")

boxcox_cf_kappa <- tidy(confusionMatrix(testing_results_kappa$diagnosis,testing_results_kappa$BoxCox)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(boxcox_cf_kappa, digits = 3, caption = "Support Vector Machine with BoxCox transformation using metric = kappa")

pca_cf_kappa <- tidy(confusionMatrix(testing_results_kappa$diagnosis,testing_results_kappa$PCA)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(pca_cf_kappa, digits = 3, caption = "Support Vector Machine with PCA using metric = kappa")

both_cf_kappa <- tidy(confusionMatrix(testing_results_kappa$diagnosis,testing_results_kappa$Both)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(both_cf_kappa, digits = 3, caption = "Support Vector Machine with both PCA and BoxCox transformation using metric = kappa")

svm_corr_cf_kappa <- tidy(confusionMatrix(testing_results_kappa$diagnosis,
                                    testing_results_kappa$SVM_corr)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(svm_corr_cf_kappa, digits = 3, caption = "Support Vector Machine without correlation using metric = kappa")


down_cf_kappa <- tidy(confusionMatrix(testing_results_kappa$diagnosis,testing_results_kappa$Down)) %>% 
  select(term:estimate) %>% 
  filter(term %in% c("specificity", "sensitivity", "f1",
                     "prevalence","recall", "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(term = str_replace(string=term, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(term = str_replace(string=term, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(term = str_replace(string=term, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(class)

kable(down_cf_kappa, digits = 3, caption = "Support Vector Machine with downsampling using metric = kappa")



```


##More model tuning

The "random" hyperparameter search which is set bu the "tunelength" argument of
the "train" function of the Caret package in R tests values for the "Cost" or
"C" hyperparameter according to a sequence of $2^X$, eg. $2^2$,$2^3$ etc.
However, values of "C" or "Cost" between eg. 4 and 8 or 16 and 24 are not tested
when assessing model performance during training on the training dataset during
cross-validation. By manually choosing values, it is possible to also test
values for the "C" or "Cost" parameter such as 6 or 7.25, that may very well be
better than 8 or 4. The same principle applies to the $\sigma$ hyperparameter,
which is chosen according to the procedure explained above and is then kept
constant during training. Thus, here I am doing a so-called "grid-search" to
test multiple combinations of different values of C and sigma.

###Tuning the best performing model

Note: The code for the manual hyperparameter tuning for the "SVM with both
BoxCox and PCA using Kappa metric" is not run when making this PDF, as the model
without the highl correlated variable proved to have a better predictive
performance overall.

Hyperparameters of the best performing  model - "SVM with both BoxCox and PCA 
using Kappa metric".

```{r parameters, eval = FALSE}

print(both_k)

plot(both_k)

```

Manually selecting values adjacent to the Cost and Sigma values identified as 
the optimal hyperparameters during the cross-validation and then training the 
model on these manually selected values for Cost and Sigma.

```{r more_tune, eval = FALSE}

grid <- expand.grid(sigma = c(0.00735,0.0074, 0.0075, 0.0076, 0.00765, 0.007,0.00725, 
                  0.0073, 0.006, 0.008, 0.0085, 0.007, 0.009, 0.01),
 C = c(6.63,6.64,6.5,6.68, 6.66, 6.625,6.675,6.69, 6.7,6.725, 6.735, 6.74, 6.75,6.775, 6.8,6.825 ,7,7.25,7.5,7.75,8, 8.25, 8.5, 8,75,9,9.25, 9.75))


set.seed(123)

svm_both_grid <- train(rec_both , data = diagnosis_train, 
                    method = "svmRadial",
                    tuneGrid = grid,
                    metric = "Kappa",
                    tuneLength = 12,
                    trControl = control)

print(svm_both_grid)

plot(svm_both_grid)


```

For the svm model with both PCA and BoxCox transformation:
Kappa was used to select the optimal model using the largest value.
The final values used for the model were sigma = 0.0075 and C = 6.64.

###Predict expanded grid model for the best performing model overall

```{r pred_tune, eval = FALSE}

testing_results_grid <- diagnosis_test %>%
        mutate("svm_grid" = predict(svm_both_grid, diagnosis_test))

SVM_grid <- conf_mat(testing_results_grid, truth = diagnosis, estimate = svm_grid)

autoplot(SVM_grid, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM using Kappa metric manual hyperparameter tuning") +
  theme(plot.title = element_text(hjust = 0.5))

SVM_grid <- (summary(SVM_grid)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(SVM_grid, digits = 3, caption = "SVM with manual hyperparameter tuning") 


```

###Tuning the second best performing model

svm on clinical, demographic and cortical thickness data with filtering of
highly correlated variables: Tuning parameter 'sigma' was held constant at a
value of 0.01128046 Kappa was used to select the optimal model using the largest
value. The final values used for the model were sigma = 0.01128046 and C = 8.



```{r params}

print(corr_k)

plot(corr_k)

```

Manually selecting values adjacent to the Cost and Sigma values identified as 
the optimal hyperparameters during the cross-validation for the second best
performing model and then training the model on these manually selected 
 values for Cost and Sigma.

```{r more_tune_two}

#The code below was used in the analysis, but due to time-saving measures,
# this code was not run in the making of this document

# grid_two <- expand.grid(sigma = c(0, 0.0075,0.0073,0.00725,0.00715, 0.007,
# 0.0095, 0.008, 0.0085, 0.006, 0.0065, 0.007, 0.009, 0.01, 0.015,0.0175, 0.02,
# 0.06,0.00675), C = c( seq(7.5,12, by = 0.1)))

#This code was not used in the initial analysis, but for illustrative purposes
#it is used for making this document.

grid_two <- expand.grid(sigma = c(0.007),
  C = c( 9))


set.seed(123)

svm_corr_grid <- train(rec_corr , data = diagnosis_train, 
                    method = "svmRadial",
                    tuneGrid = grid_two,
                    metric = "Kappa",
                    trControl = control)

print(svm_corr_grid)



```

SVM model without highly correlated variables: Kappa was used to select the
optimal model using the largest value. The final values used for the model were
sigma = 0.007 and C = 9.

###Predict expanded grid model for the second best performing model overall

results of the manually tuned SVM model which underwent the filtering of highly 
correlated variables.

```{r pred_tune_two}

testing_results_grid_two <- diagnosis_test %>%
        mutate("svm_grid" = predict(svm_corr_grid, diagnosis_test))

SVM_grid_two <- conf_mat(testing_results_grid_two, truth = diagnosis, estimate = svm_grid)

autoplot(SVM_grid_two, type = "heatmap") +
  ggtitle("Confusion matrix for the SVM model using both the cortical thicknesses\n and the clinical and the demographic data") +
  theme(plot.title = element_text(hjust = 0.5))

SVM_grid_two <- (summary(SVM_grid_two)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(SVM_grid_two, digits = 3, caption = "SVM model using both the cortical thicknesses and clinical and demographic data") 

```

Statistics for the each diagnosis:

```{r pred_tune_two_diagnosis}

manual_tuning <- tidy(confusionMatrix(testing_results_grid_two$diagnosis,
                                      testing_results_grid_two$svm_grid)) %>% 
  select("Statistic" = term, "Diagnosis" = class, "Estimate" = estimate) %>% 
  filter(Statistic %in% c("specificity", "sensitivity", "f1",
                      "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(Statistic = str_replace(string=Statistic, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(Statistic = str_replace(string=Statistic, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(Statistic = str_replace(string=Statistic, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(Diagnosis)

kable(manual_tuning, digits = 3, caption = "SVM manually tuned")


```


#SVM model based on QDEC results

SVM model based on significant areas of cortical thinning in QDEC at p < 0.05 
with monte carlo simulation for cluster-wise multiple comparisons correction.

preparing data for the training

```{r QDEC_model}

sig <- data %>% 
  select(-c(kimove_id, model))  %>% 
  select(diagnosis:h_y, lh_superiorfrontal_thickness, 
         rh_parsopercularis_thickness,rh_caudalmiddlefrontal_thickness,
         rh_precentral_thickness, rh_precuneus_thickness,
         lh_lateralorbitofrontal_thickness)

set.seed(123)

sig_split <- initial_split(sig, strata = "diagnosis", prop = .75)

sig_train <- training(sig_split)
sig_test <- testing(sig_split)

rec_sig_orig <-recipe(diagnosis ~ ., data = sig_train) %>%
step_dummy(all_nominal(), - all_outcomes())

rec_sig_svm <-recipe(diagnosis ~ ., data = sig_train) %>%
step_dummy(all_nominal(), - all_outcomes()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric())

rec_sig_down <-  recipe(diagnosis ~ ., data = sig_train) %>%
step_dummy(all_nominal(), - all_outcomes()) %>% 
step_center(all_numeric()) %>%
step_scale(all_numeric()) %>%
step_downsample(diagnosis)

rec_sig_boxcox <-recipe(diagnosis ~ ., data = sig_train) %>%
step_BoxCox(all_numeric()) %>%
step_dummy(all_nominal(), - all_outcomes()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric())

```

training models based on the QDEC results

```{r QDEC_train, eval=FALSE}

set.seed(123)

svm_sig <- train(rec_sig_svm, data = sig_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

orig_sig <- train(rec_sig_orig, data = sig_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

down_sig <- train(rec_sig_down, data = sig_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

boxcox_sig <- train(rec_sig_boxcox, data = sig_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)



```

results of the model which was based on the QDEC output

```{r QDEC_train_results, eval = FALSE}

sig_results <- sig_test %>%
        mutate("svm" = predict(svm_sig, sig_test)) %>% 
        mutate("orig" = predict(orig_sig, sig_test)) %>%
        mutate("down" = predict(down_sig, sig_test)) %>%
        mutate("BoxCox" = predict(boxcox_sig, sig_test)) 


sig_down <- conf_mat(sig_results, truth = diagnosis, estimate = down)

autoplot(sig_down, type = "heatmap") +
  ggtitle("Diagnosis Heatmap downsampled qdec results+clinical data") +
  theme(plot.title = element_text(hjust = 0.5))

sig_down <- (summary(sig_down)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(sig_down, digits = 3, caption = "downsampled qdec results+clinical data")

sig_svm <- conf_mat(sig_results, truth = diagnosis, estimate = svm)

autoplot(sig_svm, type = "heatmap") +
  ggtitle("Diagnosis Heatmap svm on qdec results+clinical data") +
  theme(plot.title = element_text(hjust = 0.5))

sig_svm <- (summary(sig_svm)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(sig_svm, digits = 3, caption = "svm on qdec results+clinical data")

sig_orig <- conf_mat(sig_results, truth = diagnosis, estimate = orig)

autoplot(sig_orig, type = "heatmap") +
  ggtitle("Diagnosis Heatmap svm without centering and scaling on qdec results+clinical data") +
  theme(plot.title = element_text(hjust = 0.5))

sig_orig <- (summary(sig_orig)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(sig_orig, digits = 3, caption = "svm without centering and scaling on qdec results+clinical data")

sig_boxcox <- conf_mat(sig_results, truth = diagnosis, estimate = BoxCox)

autoplot(sig_boxcox, type = "heatmap") +
  ggtitle("Diagnosis Heatmap BoxCox transformed qdec results+clinical data") +
  theme(plot.title = element_text(hjust = 0.5))

sig_boxcox <- (summary(sig_boxcox)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(sig_boxcox, digits = 3, caption = "BoxCox transformed qdec results+clinical data")

```


The downsampled version and the Boxcox transformed model had the best performances,
Particularly the downsampled is great at distinguishing PSP patients.

Cost and Sigma for the models using the QDEC output and clinical and demographic
data

```{r QDEC_hyperparameters, eval=FALSE}

print(svm_sig)

print(orig_sig)

print(boxcox_sig)

print(down_sig)

```

svm with scaling and centering: Tuning parameter 'sigma' was held constant at a
value of 0.1066213 Kappa was used to select the optimal model using the largest
value. The final values used for the model were sigma = 0.1066213 and C = 2.

svm without scaling and centering: Tuning parameter 'sigma' was held constant at
a value of 0.1066213 Kappa was used to select the optimal model using the
largest value. The final values used for the model were sigma = 0.1066213 and C
= 2.

BoxCox:Tuning parameter 'sigma' was held constant at a value of 0.09832538
Kappa was used to select the optimal model using the largest value.
The final values used for the model were sigma = 0.09832538 and C = 2.

Downsampling: Tuning parameter 'sigma' was held constant at a value of 0.07398176
Kappa was used to select the optimal model using the largest value.
The final values used for the model were sigma = 0.07398176 and C = 2.

##manual tuning models using qdec the results

manual tuning of the hyperparameters in the model which only used regions with
statistically significant cortical thinning in QDEC.

```{r QDEC_tuning}

# grid_sig <- expand.grid(sigma = c(seq(0.05,0.14, by = 0.001)),
#   C = c(seq(1,3.6, by = 0.1)))

grid_sig <- expand.grid(sigma = c(0.07),
  C = c(3.2))


set.seed(123)

svm_sig_grid <- train(rec_sig_down , data = sig_train, 
                    method = "svmRadial",
                    tuneGrid = grid_sig,
                    metric = "Kappa",
                    trControl = control)


print(svm_sig_grid)



```


SVM model on significant regions in QDEC with downsampling: Kappa was used to
select the optimal model using the largest value. The final values used for the
model were sigma = 0.07 and C = 3.2.

results from the manual tuning on the model based on the qdec results

```{r QDEC_final_result}

results_sig <- sig_test %>%
        mutate("down" = predict(svm_sig_grid, sig_test))

SVM_sig_down <- conf_mat(results_sig, truth = diagnosis, estimate = down)

autoplot(SVM_sig_down, type = "heatmap") +
  ggtitle("Diagnosis Heatmap QDEC significant regions") +
  theme(plot.title = element_text(hjust = 0.5))

SVM_sig_down <- (summary(SVM_sig_down)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))


kable(SVM_sig_down, digits = 3, caption = "SVM on QDEC regions+clinical data with downsampling")

```


manual tuning of the BoxCox transformed model based on the qdec results


Note: not run during the making of this PDF file, as the model with
downsampling was superior. the code below is just an example of what was done,
with no output.
 

```{r Box_qdec, eval = FALSE}

grid_box <- expand.grid(sigma = c(0.125,0.1, 0.0975),
  C = c(2.2,2.4,2.1))


set.seed(123)

svm_sig_box <- train(rec_sig_boxcox , data = sig_train, 
                    method = "svmRadial",
                    tuneGrid = grid_box,
                    metric = "Kappa",
                    trControl = control)


print(svm_sig_box)

```

significant regions model with boxcox transformation: Kappa was used to select
the optimal model using the largest value. The final values used for the model
were sigma = 0.1 and C = 2.

results from the BoxCox transformed model based on the QDEC output

```{r results_box_qdec, eval = FALSE}

results_box <- sig_test %>%
        mutate("box" = predict(svm_sig_box, sig_test))

sig_box <- conf_mat(results_box, truth = diagnosis, estimate = box)

autoplot(sig_box, type = "heatmap") +
  theme(plot.title = element_text(hjust = 0.5))

sig_box <- (summary(sig_box)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(sig_box, digits = 3)



```


#MRI only model 


Note: The code regarding the "MRI data only model" is not run when making 
this PDF document. As the conclusion states below, it this modeldid not perform
well. 


Splitting data and creating recipes for the svm models based only on mri data

```{r MRI, eval = FALSE}

mri <- data %>% 
  select(-c(kimove_id, model)) %>% 
  select(diagnosis, lh_bankssts_thickness:rh_meanthickness_thickness, 
         -lh_meanthickness_thickness)

set.seed(123)

mr_split <- initial_split(mri, strata = "diagnosis", prop = .75)

mr_train <- training(mr_split)
mr_test <- testing(mr_split)

rec_mr_both <-recipe(diagnosis ~ ., data = mr_train) %>%
step_BoxCox(all_numeric()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric()) %>%
step_pca(all_numeric(), threshold = .99)

rec_mr_corr <-recipe(diagnosis ~ ., data = mr_train) %>%
step_corr(all_numeric(), threshold = .8) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric())


rec_mr_down <-  recipe(diagnosis ~ ., data = mr_train) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric()) %>%
step_downsample(diagnosis)

rec_mr_corr_down <-  recipe(diagnosis ~ ., data = mr_train) %>%
step_corr(all_numeric(), threshold = .8) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric()) %>%
step_downsample(diagnosis)



```

training models based only on the cortical thicknesses

```{r MRI_train, eval = FALSE}

set.seed(123)

mri_both <- train(rec_mr_both, data = mr_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

mri_corr <- train(rec_mr_corr, data = mr_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

mri_down <- train(rec_mr_down, data = mr_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

mri_corr_down <- train(rec_mr_corr_down, data = mr_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

```

predicting models based on mri data only

```{r MRI_results, eval = FALSE}

mri_results <- mr_test %>%
        mutate("both" = predict(mri_both, mr_test)) %>% 
        mutate("corr" = predict(mri_corr, mr_test)) %>%
        mutate("down" = predict(mri_down, mr_test)) %>%
        mutate("corr_down" = predict(mri_corr_down, mr_test)) 

mr_both <- conf_mat(mri_results, truth = diagnosis, estimate = both)

autoplot(mr_both, type = "heatmap") +
  ggtitle("Diagnosis Heatmap mri data only with both boxcox transformation and pca") +
  theme(plot.title = element_text(hjust = 0.5))

mr_both <- (summary(mr_both)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(mr_both, digits = 3, caption = "mri data only with both boxcox transformation and pca")

mr_corr <- conf_mat(mri_results, truth = diagnosis, estimate = corr)

autoplot(mr_corr, type = "heatmap") +
  ggtitle("Diagnosis Heatmap MRI data only without high correlations") +
  theme(plot.title = element_text(hjust = 0.5))

mr_corr <- (summary(mr_corr)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(mr_corr, digits = 3, caption = "mri data only without highly correlated variables")

mr_down <- conf_mat(mri_results, truth = diagnosis, estimate = down)

autoplot(mr_down, type = "heatmap") +
  ggtitle("Diagnosis Heatmap SVM using mri only data with downsampling") +
  theme(plot.title = element_text(hjust = 0.5))

mr_down <- (summary(mr_down)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(mr_down, digits = 3, caption = "mri data only With downsampling")

mr_corr_down <- conf_mat(mri_results, truth = diagnosis, estimate = corr_down)

autoplot(mr_corr_down, type = "heatmap") +
  ggtitle("Diagnosis Heatmap MRI data only with downsampling and without correlations") +
  theme(plot.title = element_text(hjust = 0.5))

mr_corr_down <- (summary(mr_corr_down)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(mr_corr_down, digits = 3, caption = "MRI data only without highly correlated variables and with downsampling")

```

##Conclusion from using MRI data only

Using only the cortical thicknesses does not seem to predict the diagnoses well.
The manual hyperparameter tuning was not attempted for the model based on the
MRI data, as it was not considered to add any value due to the models initial 
poor performance.

#Model using solely clinical data

Demographic and clinical data: age, gender, Hoehn and Yahr score, symptom 
duration


```{r clin_split}

clin <- data %>% 
  select(-c(kimove_id, model)) %>% 
  select(diagnosis:h_y)

set.seed(123)

clin_split <- initial_split(clin, strata = "diagnosis", prop = .75)

clin_train <- training(clin_split)
clin_test <- testing(clin_split)

rec_clin_svm <-recipe(diagnosis ~ ., data = clin_train) %>%
step_dummy(all_nominal(), - all_outcomes()) %>%
step_center(all_numeric()) %>%
step_scale(all_numeric())

rec_clin_down <-  recipe(diagnosis ~ ., data = clin_train) %>%
step_dummy(all_nominal(), - all_outcomes()) %>% 
step_center(all_numeric()) %>%
step_scale(all_numeric()) %>%
step_downsample(diagnosis)

```

training


Note: the code in the chunk below was not run. The next chunk 
includes the manual
optimization of the model which onl uses clinical data, which is the final
result for the model based on only the clinical and demographic data.


```{r clin_training, eval = FALSE}

set.seed(123)

svm_clin <- train(rec_clin_svm, data = clin_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

set.seed(123)

down_clin <- train(rec_clin_down, data = clin_train,
method = "svmRadial",
metric = "Kappa",
tuneLength = 12,
trControl = control)

testing_clin <- clin_test %>%
        mutate("svm" = predict(svm_clin, clin_test)) %>% 
        mutate("down" = predict(down_clin, clin_test))

SVM_clin <- conf_mat(testing_clin, truth = diagnosis, estimate = svm)

autoplot(SVM_clin, type = "heatmap") +
  theme(plot.title = element_text(hjust = 0.5))

SVM_clin <- (summary(SVM_clin)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(SVM_clin, digits = 3) 

Down_clin <- conf_mat(testing_clin, truth = diagnosis, estimate = down)

autoplot(Down_clin, type = "heatmap") +
  theme(plot.title = element_text(hjust = 0.5))

Down_clin <- (summary(Down_clin)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))

kable(Down_clin,digits = 3)


```


Svm only: Tuning parameter 'sigma' was held constant at a value of 0.3903809
Kappa was used to select the optimal model using the largest value.
The final values used for the model were sigma = 0.3903809 and C = 2.

Downsampling: Tuning parameter 'sigma' was held constant at a value of 0.1754092
Kappa was used to select the optimal model using the largest value.
The final values used for the model were sigma = 0.196 and C = 3.


manual selection of tuning parameters of the clinical data only

```{r clin_tune}



grid_two <- expand.grid(sigma = c(0.196),
  C = c( 3))

set.seed(123)

svm_clin_grid <- train(rec_clin_down , data = clin_train, 
                    method = "svmRadial",
                    tuneGrid = grid_two,
                    metric = "Kappa",
                    tuneLength = 12, 
                    trControl = control)

print(svm_clin_grid)


```

test the model based on clinical and demographic data

```{r clin_tune_results}

testing_clin <- clin_test %>%
        mutate("down" = predict(svm_clin_grid, clin_test))

Down_clin <- conf_mat(testing_clin, truth = diagnosis, estimate = down)

autoplot(Down_clin, type = "heatmap") +
  ggtitle("Confusion matrix for the SVM model using only the clinical\n and the demographic data") +
  theme(plot.title = element_text(hjust = 0.5))

Down_clin <- (summary(Down_clin)) %>% 
  select("Metric" = .metric, "Estimate" = .estimate) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='f_meas', replacement='F-score')) %>% 
 mutate(Metric = str_replace(string=Metric, pattern='j_index', replacement="Youden's index")) %>% 
  mutate(Metric = str_replace(string=Metric, pattern='kap', 
                              replacement="kappa")) %>% 
  filter(Metric %in% c("specificity", "sensitivity", "F-score", 
                       "Youden's index", "accuracy", "kappa", "precision",                              "recall"))


kable(Down_clin, digits = 3, caption = "SVM on clinical data only with downsampling") 

```

results by diagnosis model, using clinical variables only.


```{r clin_by_diagnosis}

clinical <- tidy(confusionMatrix(testing_clin$diagnosis,
                                      testing_clin$down)) %>% 
  select("Statistic" = term, "Diagnosis" = class, "Estimate" = estimate) %>% 
  filter(Statistic %in% c("specificity", "sensitivity", "f1",
                      "pos_pred_value", 
                     "neg_pred_value" )) %>%
  mutate(Statistic = str_replace(string=Statistic, pattern='f1', 
                            replacement='F-score')) %>% 
  mutate(Statistic = str_replace(string=Statistic, pattern='pos_pred_value', 
                            replacement='ppv')) %>%  
  mutate(Statistic = str_replace(string=Statistic, pattern='neg_pred_value', 
                            replacement='npv')) %>%  
  arrange(Diagnosis)


kable(clinical, digits = 3, caption = "SVM on clinical data only with downsampling") 
  


```


#Summary

The model with the best performance was SVM done on the whole data which first 
underwent a filtering by correlation using a threshold of 0.8 for the Pearson
Correlation Coefficient. It can be added that this threshold was used following 
extensive examination of multiple threshold values.The data was then scaled and 
centered. The model was initially run alongside multiple other models, 
with the hyperparameters Cost and Sigma being tuned automatically during a 12-fold Cross-validation repeated 12 times. The model which had undergone this removal 
of variables by correlation exhibited the second best performance among all
models after testing on the testdataset. The best model at that time was the one 
in which initially a Box-Cox transformation was done, and then the variables 
were scaled, centered and runthrough a PCA and then the Principle Components 
explaining 99% of the variancewere used in making that particular model. 

The two best models from the initial testing were then used in yet another 
training phase, but this time the hyperparameters were set manually based on 
values that are located adjacent to the Cost and Sigma parameters that were 
chosen during the Cross-validation. After continous trial and error, the 
model that underwent removal of variables based on correlation emerged as the
model which had the best performance. All columns from the original dataset, 
except for "Kimove ID", were used in creating this model. Although we suspected 
that e.g. the total intracranial volume or Hoehn and Yahr would not contribute 
to the model, it was eventually showed that the model performed best when all
the columns were used. 

The parameters of the final model were as follows:
sigma = 0.007 and C = 9. 

A total of 16 variables were excluded from this model during pre-processing,
based on which ones exhibited correlation coefficients greater than 0.8 in 
relation to another variable in the dataset.

#initial SVM results table

##Collecting all model predictions into a single object

```{r final}



Orig_k$model <- as.vector(rep("Orig_k", nrow(Orig_k)))

PCA_k$model <- as.vector(rep("PCA_k", nrow(Orig_k)))

Down_k$model <- as.vector(rep("Down_k", nrow(Orig_k)))

Both_k$model <- as.vector(rep("Both_k", nrow(Orig_k)))

BOXCOX_k$model <- as.vector(rep("BOXCOX_k", nrow(Orig_k)))

SVM_k$model <- as.vector(rep("SVM_k", nrow(Orig_k)))

SVM_corr_k$model <- as.vector(rep("Corr_k", nrow(Orig_k)))

Orig$model <- as.vector(rep("orig", nrow(Orig_k)))

PCA$model <- as.vector(rep("pca", nrow(Orig_k)))

Down$model <-  as.vector(rep("down", nrow(Orig_k)))

Both$model <-  as.vector(rep("both", nrow(Orig_k)))

BOXCOX$model <-  as.vector(rep("boxcox", nrow(Orig_k)))

SVM$model <-  as.vector(rep("svm", nrow(Orig_k)))

SVM_corr$model  <-  as.vector(rep("corr", nrow(Orig_k)))


final <- bind_rows(Orig_k, BOXCOX_k, SVM_corr_k, SVM_k,  PCA_k, Both_k, Down_k, Down, SVM_corr, Orig,  SVM, BOXCOX, PCA, Both)


```



## Initial model training results table


```{r final_table}

options(knitr.kable.NA = '0')

final %>%  
  filter(Metric %in% c("F-score", "kappa", "Youden's index")) %>% 
  mutate(model = str_replace(string=model, pattern='svm', replacement="SVM on centered and scaled data")) %>% 
  mutate(model = str_replace(string=model, pattern='down', replacement="SVM on downsampled data"))%>% 
 mutate(model = str_replace(string=model, pattern='corr', replacement="SVM on data without highly correlated variables"))%>% 
 mutate(model = str_replace(string=model, pattern='boxcox', replacement="SVM on data with BoxCox transformation"))%>% 
 mutate(model = str_replace(string=model, pattern='pca', replacement="SVM on principle components")) %>% 
  mutate(model = str_replace(string=model, pattern='both', replacement="SVM on data with both boxcox and pca")) %>% 
  mutate(model = str_replace(string=model, pattern='orig', replacement="SVM on unprocessed data")) %>% 
    mutate(model = str_replace(string=model, pattern='cory', replacement='SVM without high correlation and on principle components using kappa metric')) %>% 
 mutate(model = str_replace(string=model, pattern='Orig_k', replacement="SVM on unprocessed data using kappa metric")) %>% 
 mutate(model = str_replace(string=model, pattern='BOXCOX_k', replacement="SVM on data with BoxCox transformation using metric kappa")) %>% 
 mutate(model = str_replace(string=model, pattern='Corr_k', replacement="SVM on data without correlated variables using metric kappa"))%>% 
 mutate(model = str_replace(string=model, pattern='PCA_k', replacement="SVM on principle components using kappa metric"))%>% 
 mutate(model = str_replace(string=model, pattern='Both_k', replacement="SVM on boxcoxed and PCA data using kappa metric"))%>% 
 mutate(model = str_replace(string=model, pattern='Down_k', replacement="SVM on downsampled data using kappa metric"))%>% 
 mutate(model = str_replace(string=model, pattern='SVM_k', replacement="SVM on centered and scaled data using kappa metric ")) %>% 
  kable(digits = 3, caption = "Model summary") 
  


```

#Session info

```{r session}

sessioninfo::session_info()

```

